Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2011-11-18 Multiresolution_framework/ltx/manuscript.tex
added subplots

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
\end{align}
The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}\label{eq:CovMatrix}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}
\section{B-spline MRAIDE in State-Space }
B-spline functions are one of the most basic elements for wavelet construction, these functions are composed of  piecewise polynomials combined smoothly at some breaking points (knots) which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \cite{Goswami1999} where the higher the order, the smoother the resulting B-spline becomes (see Figure~\ref{fig:Figure0}).
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure0a.pdf}
\includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (coloured lines) are joined at the break points (black dots). (\textbf{a}) B-spline function of order 2 (linear B-spline). (\textbf{b}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

@@

To determine the level of approximation in the model, detailed in the previous section, that is the number of basis functions required to reconstruct the field, $v_t(r)$, the frequency response of the B-spline function needs to be computed by applying the Fourier transform to \eqref{eq:N1convolutions}. To proceed the Fourier transform of $N_1(r)$ is first calculated, 
\begin{align}\label{eq:MRA-N1Fouriertransform}
\mathcal F(N_1(r))&=\int_{-\infty}^{+\infty}N_1(r)\mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\int_{0}^{1} \mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}.
\end{align}
The convolution theorem states that the Fourier transform of the convolution of functions is equal to the product of their Fourier transforms. Therefore, taking Fourier transform of \eqref{eq:N1convolutions} and substituting \eqref{eq:MRA-N1Fouriertransform} for $\mathcal F(N_1(r)) $ yields
\begin{align}\label{eq:MRA-NmFouriertransform}
\mathcal F(N_m(r))=\left(\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}\right)^m.
\end{align}
The multi-resolution approximation using B-spline functions is completed by defining a two-scale relation pair which for cardinal B-spline scaling and wavelet functions of order $m$, takes the form of \cite{Chui1992}
\begin{align}
 N_{m}\left(r\right)&=\sum_{n=0}^{m} \left[\mathbf p\right]_n N_{m}\left(2r-n\right) \label{eq:MRA-TwoScalepair1} \\
  \varphi_{m}\left(r\right) &= \sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m-1}} \left[\mathbf q\right]_n N_{m}\left(2r-n\right)\label{eq:MRA-TwoScalepair2},
 \end{align}
 where $\left[\cdot\right]_n$ denotes the $n$-th element of the vector coefficients, and
 \begin{align}
\left[\mathbf p\right]_n&=2^{-m+1} \binom{m}{n} \quad \text{ $0\le n\le m$} \label{eq:MRA-TwoScalepair1coefs}\\
\left[\mathbf q\right]_n&= \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$}\label{eq:MRA-TwoScalepair2coefs}.
 \end{align}
 Figure~\ref{fig:MRA-Figure1} shows the cubic B-spline scaling function and its associated wavelet function with dilation and translation parameters set to zero. 
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure1a.pdf}
\includegraphics{./Graph/Figure1b.pdf}
\caption{{\bf Examples of B-spline  scaling and wavelet functions}. (\textbf{a}) The cubic B-spline scaling function. (\textbf{b}) The corresponding wavelet function.}
\label{fig:MRA-Figure1}
\end{figure}

@@

 		\includegraphics[scale=1]{./Graph/ScaleDecomposition.pdf}\\
 		\includegraphics[scale=1]{./Graph/WaveletDecomposition.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. The scaling and the wavelet functions (blue) at $j=0$ are respectively obtained using 625 and 1250 scaling functions (black) at $j=4$.}

@@

 In this paper, 4th order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating \eqref{eq:Lambdax}, \eqref{eq:Lambdatheta} and \eqref{eq:CovMatrix} whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.

>>>>

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
\end{align}
The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}\label{eq:CovMatrix}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}
\section{B-spline MRAIDE in State-Space }
B-spline functions are one of the most basic elements for wavelet construction, these functions are composed of  piecewise polynomials combined smoothly at some breaking points (knots) which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \cite{Goswami1999} where the higher the order, the smoother the resulting B-spline becomes (see Figure~\ref{fig:Figure0}).
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure0.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (coloured lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

@@

To determine the level of approximation in the model, detailed in the previous section, that is the number of basis functions required to reconstruct the field, $v_t(r)$, the frequency response of the B-spline function needs to be computed by applying the Fourier transform to \eqref{eq:N1convolutions}. To proceed the Fourier transform of $N_1(r)$ is first calculated, 
\begin{align}\label{eq:MRA-N1Fouriertransform}
\mathcal F(N_1(r))&=\int_{-\infty}^{+\infty}N_1(r)\mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\int_{0}^{1} \mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}.
\end{align}
The convolution theorem states that the Fourier transform of the convolution of functions is equal to the product of their Fourier transforms. Therefore, taking Fourier transform of \eqref{eq:N1convolutions} and substituting \eqref{eq:MRA-N1Fouriertransform} for $\mathcal F(N_1(r)) $ yields
\begin{align}\label{eq:MRA-NmFouriertransform}
\mathcal F(N_m(r))=\left(\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}\right)^m.
\end{align}
The multi-resolution approximation using B-spline functions is completed by defining a two-scale relation pair which for cardinal B-spline scaling and wavelet functions of order $m$, takes the form of \cite{Chui1992}
\begin{align}
 N_{m}\left(r\right)&=\sum_{n=0}^{m} \left[\mathbf p\right]_n N_{m}\left(2r-n\right) \label{eq:MRA-TwoScalepair1} \\
  \varphi_{m}\left(r\right) &= \sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m-1}} \left[\mathbf q\right]_n N_{m}\left(2r-n\right)\label{eq:MRA-TwoScalepair2},
 \end{align}
 where $\left[\cdot\right]_n$ denotes the $n$-th element of the vector coefficients, and
 \begin{align}
\left[\mathbf p\right]_n&=2^{-m+1} \binom{m}{n} \quad \text{ $0\le n\le m$} \label{eq:MRA-TwoScalepair1coefs}\\
\left[\mathbf q\right]_n&= \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$}\label{eq:MRA-TwoScalepair2coefs}.
 \end{align}
 Figure~\ref{fig:MRA-Figure1} shows the cubic B-spline scaling function and its associated wavelet function with dilation and translation parameters set to zero. 
\begin{figure}[!ht]
\centering
\includegraphics{./Graph/Figure1.pdf}
% \includegraphics{./Graph/Figure1b.pdf}
\caption{{\bf Examples of B-spline scaling and wavelet functions}. (\textbf{A}) The cubic B-spline scaling function. (\textbf{B}) The corresponding wavelet function.}
\label{fig:MRA-Figure1}
\end{figure}

@@

 		\includegraphics[scale=1]{./Graph/Decomposition.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (blue) at $j=0$ is obtained using 625 scaling functions (black) at $j=4$. (\textbf{B}) The wavelet function (blue) at $j=0$ is obtained using 1250 scaling functions (black) at $j=4$.}

@@

 In this paper, 4th -- order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating \eqref{eq:Lambdax}, \eqref{eq:Lambdatheta} and \eqref{eq:CovMatrix} whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.
